### 2019-01-30 texs/paper_base.tex
mazu teisai

\documentclass[senior,final,11pt]{iscs-thesis}
%論文の種類とフォントサイズをオプションに
%\usepackage{graphicx}% 必要に応じて
%\usepackage{mysettings}% 自分用設定
%-------------------
\etitle{Neural Network based Decompiler for automatically recovering source codes from compiled binary data}
\jtitle{ニューラルネットを用いた逆コンパイラ}

@@

\usepackage{listings}
\usepackage{cite}
\usepackage{url}
\usepackage{natbib}
\usepackage[dvipdfmx]{graphicx}
\usepackage{framed}
\usepackage{caption}

@@

\begin{document}
\begin{eabstract}
A decompiler is a tool for recovering a source code from compiled binary data.
There are various decompilers, but they often output a source code that is not intelligible to humans. 
In this thesis, we apply machine translation techniques to generate human intelligible decompiled source codes. 
More specifically, we propose to use recurrent neural networks with attention, 
which have been demonstrated to be useful for machine translation. 
In experiments, we use source codes collected from open source projects and their binary data for training and evaluating the neural networks.
From the experiment, we can observe little improvement in the decompilation result with the use of attention.
However, the results were not applicable for practical use.

@@

In this thesis, we apply neural networks to the problem of decompilation to generate human-friendly pseudocodes. 
This idea is originated from \cite{Motoneta}, in which they used recurrent neural networks, in particular sequence-to-sequence model, for decompilation.
However, their model does not seem to have some features for improving the quality of decompilation.
Therefore, we implemented the attention mechanisme, which is introduced in \cite{dot_attention}, for improving the quality.
Additionaly, the sequence-to-sequence model, which they used for modeling, ignores the structure of the C language code.
Therefore, we also implemented the sequence-to-tree model, which is suggested in \cite{Seq2Tree}, for generating structured C language source code.

In the experiment, we tested three models, the basic sequence-to-sequence model, the sequence-to-sequence model with attention and the sequence-to-tree model.
As the result of the experiment, the attention mechanism showed little improvemtn in the performance, while the sequence-to-tree does not.
Unfortunately, the results were not applicable for practical use, because the most of decompilation results were incorrect.

@@

\chapter{Related Works}
In this section, we review related works.

@@

In \cite{Motoneta}, they used the seqence-to-sequence model, which is a popular technique in the machine translation.
The seqence-to-sequence model is also evaluated in this thesis.

In \cite{genetic_decompiler}, they applied genetic programming to decompilation.
They prepared big corpus of C source codes, and they repeatedly mutated C code to reduce the difference of the target binary and the compilation result of the source code.

@@

\section{Improvement of decompilation results}
In this thesis, we do not try to recover the function name and variable name. 
However, for hand-decompilation, sometimes the name can be estimated from the functionality of the variable.
For example, \cite{hand_decompilation} used Boomerang decompiler and repeatedly fixed variable types and names of the decompilation results 
to recover the source code of a software.

@@

\cite{name_recover_from_decompile_result} tackled this work. 
In that paper, they tried to estimate the name of a variable which appears in the decompilation result of a compiled C program.

This kind of name estimation also appears in the context of deobfuscation.
\cite{deobfsucation_matome}
% これ雑にreferenceしたけどあんま意味ないよなぁ...?
Deobfuscation is a reverse process of obfuscation.
Obfuscation is a technique to modify the program source code so that the code becomes hard to understand for humans. 
Obfuscation is used for hiding the behavior of a program whose source code has to be distributed.
For example, in the web application, a JavaScript code of a program should be distributed in order to be executed on client web browsers.
To protect the program from cracking, the JavaScript code is sometimes minified. All comments, newlines, indents are removed and variable names are substituted to automatically generated meaningless names.

\cite{JSNaughty} applied statistical machine translation for this task. 
They calculated the future value of a variable from its functionality, then 
sought the most suitable name from the candidates generated from the corpus.

@@

There are many applications of neural networks for program synthesis. They are summurized in \cite{deep_programming_matome}.

In this thesis, we applied sequence-to-tree model suggested in \cite{Seq2Tree}, because the input of the decompiler is the sequence of machine codes.

For another example, in \cite{coffeescript_to_javascript}, they tried to mutually translate from JavaScript to CoffeeScript, and its inverse.
In that case, both JavaScript codes and CoffeeScript codes are tree structured, so they suggested tree-to-tree model.

\chapter{Problem Settings and Methots}

@@

A compiler is a program which compiles a source code written in language X to another language Y. 
% 文献 Usually, X is a high-level language and Y is relatively a low-level language. 
For example, a GNU C compiler translates the C language to the x64 assembly language, a javac compiler translates Java to a Java virtual machine code.
% This process is usually 
% 文献
% Especially, 
% This reverse process is usually insufficient or impossible.
A decompiler is a program which aims to reverse the process of a compiler, retrieving a source code of the high-level language X from a source code of a low-level language Y. 
When X is a high-level language and Y is relatively a low-level language, 
this reverse process becomes insufficient or impossible
\citep{hex_rays,decompile_hard_java}.
% for instance X is the C language and Y is the x64 assembly language 
This is because a lot of informaition, such as variable names or function names, are lost when compiling.
Additionally, the program structure described in the high level-language is lost too. 
For example, a simple \texttt{for} statement can be represented by the combination of a \texttt{while} statement and an \texttt{if} statement, or the combination of \texttt{if} statement and \texttt{goto} statement. 
So a decompiler cannot distinguish the difference in their representations, 
although in the real situation \texttt{for} statement are often used and \texttt{goto} statement are rarely used.

@@

By utilizing a corpus for soruce codes, a statistical method might have the potentioal for finding patterns automatically, without enumerating each patterns.
Therefore, we tried to apply some statistical methods, which are described in the next section.

@@

% 189
In this thesis, we try to apply statistical approach, specifically SMT (statistical machine translation) technique, for decompilation.
SMT is the task which tries to translate one sentense $sx$ written in one language into the sentense $sy$ written in another language.

The decompilation problem is formulated as follows. 
Suppose that there is a compiler $ f $ which converts a source code $ sx $ written in a programming language 
to a source code $ sy $  written in another programming language Y with probability $p(sy|sx)$.
In addition, suppose that the human-generated source code $ sx $ follows the probability distrubution $ p(sx) $.
In this situation, the decompilation probability distribution $ p(sx|sy) $ in proportion to $ p(sy|sx)p(sx) $ according to the Bayesian low, 
and the $\argmax_{sx} p(sx|sy)$ is considered as the best human-intelligible decompilation result for the code $sy$.

In the experiment, we collected source codes from open source projects as the random sampling from the distribution $ p(sx) $, 
and compileded them to generate the data pairs $ {\{(sx_k,sy_k)\}}_k $.
Then the data pairs are used for training the SMT models to estimate $p(sx|sy)$, and testing the performance of the models.

@@

% (TODO)LSTM,Seq2seqの説明、その後Attentionについて説明、さらにその後seq2treeについて説明する。
\subsection{RNN and LSTM}
The RNN (Recurrent Neural Network) is a neural network for dealing with sequence data.
The structure of the RNN is the sequence of the same node units, as shown in figre~\ref{fig:encoder} and figre~\ref{fig:decoder}. 
The LSTM (Long Short-Term Memory) is a unit of Recurrent Neural Network. 
It was first introduced in \cite{first_LSTM}, for overcoming the vanishing gradient problem and exploding gradient problem.
In short, LSTM unit recieves two input vectors, 
the one is the internal state of the LSTM, and the other is the input for the LSTM unit, 
and LSTM emits one output vectors, which represents the next internal state of the LSTM.

@@

The sequence-to-sequence (sequence to sequence) model is the model for SMT task, introduced in  \cite{seq2seq}.
% The combination of RNN and LSTM is applied to SMT task in, as the essential unit of the 
Sequence-to-sequence model is composed of two recurrent neural networks, the encoder netwrok and the decoder network.

@@

Figure figre~\ref{fig:encoder} shows the strucure of the encoder network.  The purpose of the encoder network is to summerize the input sequence $sx$, whose length varies depending on the sequence, into one fixed length vector.
This process is carried out as follows. 
First, $sx$ is tokenized to sequence $sx_{1} \dots sx_{n}$, and each token $sx_{k}$ is embedded to its vector representation $vx_{k}$. 
Then, the initial LSTM internal state $enc_0$ is intialized randomely, 
and $ enc_{1} \dots enc_{n} $ are calculated in accordance with the recursive formula $ enc_{k} = lstm_{enc}(enc_{k-1},vx_{k}) $.
Finally, the last internal state $ enc_{n}$ is returned as the reduced representation of the input sequence $sx$.

Figure figre~\ref{fig:decoder} shows the strucure of the decoder network.  The purpose of the decoder network is to estimate the distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $ for each k. 
The training process is carried out as follows. 
First, $sy$ is tokenized, embedded to a vector representation $vy_{1} \dots vy_{m}$ in the same way as embedding $sx$.
% prepended the token $ EOS $ for indicating the end of the sequence, then 
Then, the first internal state $ dec_{1}$ is calcuated from the start vector $vy_{begin}$ and encoded input $ enc_{n} = dec_{0} $ in accordance with the formula  $ LSTM_{dec}(enc_{n},y_{begin}) $.
Then, $ dec_{2} \dots dec_{m+1} $ are calculated in accordance with the recursive formula $ dec_{k+1} = lstm_{dec}(dec_{k},vy_{k}) $.
Each decoder internal state $ dec_{k} $ is passed to the linear layer $ W_{dec} $ and then softmaxed to generate a probability vector $ py_{k} $.
The probability vector $ py_{k} $ is intended to represent the probability distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
Thus the network is trained to reduce the difference between predicted distribution $ py_{k} $ and actual distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
In the experiment, we choiced cross-entropy of $ py_{k} $ and $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $ for loss function.

@@

First is the Bidirectional Reccurent Neural Networks (BRNN) is the architecture introduced in \cite{BiRNN}.
In the experiment, this architecture is used for the encoder network as follows.
In addition to the internal state $ enc_{k} $, the reverse internal state $ enc_{k}^{\gets} $ is also calculated 
in accordance with the recursive formula $ enc_{k}^{\gets} = lstm_{enc}^{\gets}(enc_{k+1}^{\gets},vx_{k}) $.
Then both of the internal states $ enc_{n} $ and $ enc_{1}^{\gets} $ are used for decoder network.

Second is the multi-layer reccurent neural networks, introduced in \cite{multi_layer}.
They reported that, feeding the output sequence of the LSTM network to another LSTM network, like stucking the network layers, will improve the accuracy of the prediction.
This architecture is called as stucked reccurent neural networks or multi-layer reccurent neural networks.

@@

The attention module is first developed by \cite{attention_paper}.
We inplemented the global general attention, which is intoduced in \cite{dot_attention}, for improving the quality of decompilation.
In the following, we will describe about the global general attention.

As above mentioned, the pure sequence-to-sequence model uses the output of decoder network $ dec_{k} $ for estimating 
the distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
In the sequence-to-sequence model with attention, 
a context vector $ ctx_{k} $ is calculalted from $ dec_{k} $ and the internal states of encoder network $ enc_{1} \dots enc_{n} $, 
then the distribution is estimated from the output vector $ dec_{k} $ and context vector $ ctx_{k} $.
The context vector $ ctx_{k} $ is calculated as follows.
For each state $ enc_{l} $, the importance of the state $ a_{l} $ is calculalted as $ a_{l} = score(dec_{k},enc_{l}) $,
then the $ ctx_{k} $ is calculated as $ ctx_{k} = \sum_{l=1}^{n} \frac{\exp{a_{l}}}{\sum_{l=1}^{n}\exp{a_{l}}} enc{l} $, as the weighted average of $ enc_{1} \dots enc_{n} $.
% Multiple score functions are introduced by \cite{dot_attention}, and we choiced 
In the paper by \cite{dot_attention}, various kinds of score function is suggested.
We used general score function, which is calculated as $ score(dec_{k},enc_{l}) = dec_{k}^T W enc_{l} $ ,
where W is the parameter matrix.

@@

\subsection{Sequence-to-Tree}
% 184 words
Compare to natural languages, programming languages obeys rigid grammers. % wich are usually context free grammers.
Therefore, it is reasonable to consider the grammer of the programming language when generating programs.
Sequence-to-Tree is a model for generating tree-structured output from sequence input. 
This model is introduced in \cite{Seq2Tree} for generating a program from a program specification written in the natural language. 
This task is similer to our task, which aims to generate programs from the sequence of the assembly code, so we implemented this model in the experiment.

Now, we will describe about the sequence-2-tree model.
sequence-to-tree model is composed of two modules, the encoder netwrok and the decoder network.
The encoder network of sequence-2-tree model is the same as the encoder network of sequence-to-tree model. It summerize the input sequence into one fixed length vector.

@@

The decoder network is a network for generating tree structured output. 
A tree-structure is representable by the sequence of the actions which expands a nonterminal symbol of the constructing tree.
For example, the tree shown in figre~\ref{fig:ast_zu} is gererated from the sequence $ [act_1, \dots act_{12}] $, 
where action $ act_k $ expands the left-most nonterminal symbol in the tree generated from the action sequence $ [act_1, \dots act_{k-1}] $. 
Suppose the tree $ty$ is generated by the action sequence $ [act_1, \dots act_m] $, 
then the tree-generating probability $ p(ty|sx) $ is decomposed to $ \prod_{k=1}^n p(act_k|sx,[act_1, \dots act_{k-1}]) $. 
The sequence-to-sequence model is now applicable for generating tree structure, by emmbeding each action to representing vector $av_k$ and 
calculating $ LSTM_{dec}(dec_{k},av_{k}) $ sequentially for estimating $p(act_{k+1}|sx,[act_1, \dots act_{k}]) $.

@@

In that paper, they additionaly tries to use the information which is peculiar to the tree structure. 
For expanding the node $d$, they also use the information of node type $ nf_{d} $ and the parent action $p_{d}$ from which $d$ is generated.
And they calculalte $ LSTM_{dec}(dec_{k},[av_{k}; nfv_{d}; pv_{d}]) $ instead of $ LSTM_{dec}(dec_{k},av_{k}) $, 
where $nfv_{d},pv_{d}$ are the vector embedding of $nf_{d},p_{d}$.

@@

\chapter{Experiments}
In this chapter, we explain the detail of the data preparetion and implementation.
The implementation source codes are avaiable at \url{https://github.com/satos---jp/neural_decompiler}.

\section{Data Collection}
% 219 words
We collected source codes by crawling \cite[GitHub]{github}, cloning about 900 repositories with C language topic, 
and recursively searched each folder to enumerate C language source code.
From each source code, we generated pairs of C source code fragment and x64 assembly code corresponds to the C fragment.
Some example data pairs are shown in figure~\ref{fig:pairsoffragments}.

@@

The pairs were generated as follows. Let {\sl S0} be an original C source code.
\begin{enumerate}
\item Remove all preprocessor, such as {\sl \#include}, in {\sl S0} to generate preprosessed C code {\sl S1}. 
\item Tokenize {\sl S1} and rearrange them in {\sl S2} so that there is exactly one token for each line. 
\item 
Compile {\sl S2} by gcc (GNU C compiler) with no optimization (-O0) and debug option to get assembly code {\sl S3}. 
With debug option, {\sl S3} contains information about the source code position which each assembly operations are originated from.
\item Assemble {\sl S3} to object file {\sl S4} and disassemble {\sl S4} to get information eliminated assembly code {\sl S5}.
\item 
Parse {\sl S2} with clang to get AST (abstract syntax tree) of the source code. 
And for each subtree of AST, we can get C fragment source code which corresponds to the subtree, 
and we can get assembly code corresponds to the C fragment by employing {\sl S5} combine with {\sl S3}.
\end{enumerate}

@@

Second way is the tokenized sequence of assembly codes. 
The disassembled opcodes and operands in {\sl S5} are split to symbols, and concatenated these sequences to one sequence, then we modified followings.
\begin{itemize}
\item For each end of the operations, the {\sl ENDLINE} symbol is added to indicate the end of operation.
\item 
The relative access of code address, such as {\sl call} or {\sl jump} operations, 
are relabeled like $\alpha$-conversion for making code position independent and for reducing the vocabulary size.
\item The big integer values are substituted. 
In detail, the integer in the range from -255 to 256 are preserved as it is, and the other integeres are substituted with the {\sl \_\_HEX\_\_} symbol.
This also reduces the vocabulary size.
\end{itemize}

We choiced second way for the experiment.
This is because, in our case, the assembly code is generated from common C compiler, therefore the assembly code can be disassembled uniquely.
For example, \cite{disasm_obfuscate} proposed a obfuscation technique, which deceives the disassembler to generate wrong disassembly result.
In that case, the first way might be better than the second way.

\subsection{C Code Preprocess}
% 232 words
The C source codes were parsed to AST by the \cite{clang} and its Python bindings together with \cite{pycparser}, 
because all of the tools are individualy not sufficient for generating dataset.
Then, the AST is converted to input data by following modifications.

\begin{itemize}
\item The variable-ary nodes and optional nodes are converted. 
For example, the {\sl CompoundStatement} node has variable number of statements. 
We converted such list of nodes to the chains of concatnation {\sl CONS} and termination {\sl NIL}, like a list representation in functional language.
Another example is the  {\sl If} statement, whose {\sl else} clausure are sometimes missing. 
We converted such optional node to the alternative of {\sl SOME} and {\sl NONE}, like a optional data structure in functional language.
\item The variable names, function names, label names, or type names were renamed like $\alpha$-conversion.
\item There are four type of literals in C language. 
The integer literal less than 256 is preserved, and the literal greater than 256 is clipped to 256.
The string, float and character literals are converted to {\sl \_\_LITERAL\_\_} simbol.

Modifying the node to indicate the value in the assembly code corresponding to the literal is also conceibable way,
but we did not implement this because of its technical difficalty, so this is left to future work.
\end{itemize}

@@

\section{Model Configuration}
The training model was implemented with Chainer version 4.5.0. (\url{https://chainer.org/})
We implemented three models, Seq2seq, Seq2seqAtt, and Seq2tree.

>>>>

\documentclass[senior,final,11pt]{iscs-thesis}
%論文の種類とフォントサイズをオプションに
%\usepackage{graphicx}% 必要に応じて
%\usepackage{mysettings}% 自分用設定
%-------------------
\etitle{Neural Network based Decompiler for automatically recovering source codes from compiled binary data}
\jtitle{ニューラルネットを用いた逆コンパイラに}

@@

\usepackage{listings}
\usepackage{cite}
\usepackage{url}
\usepackage[numbers]{natbib}
\usepackage[dvipdfmx]{graphicx}
\usepackage{framed}
\usepackage{caption}

@@

\begin{document}
\begin{eabstract}
A decompiler is a tool for recovering a source code from compiled binary data.
There are various decompilers, but they often output a source code that is not intelligible to humans. 
In this thesis, we apply machine translation techniques to generate human intelligible decompiled source codes. 
More specifically, we propose to use recurrent neural networks with attention, 
which have been demonstrated to be useful for machine translation. 
In experiments, we use source codes collected from open source projects and their binary data for training and evaluating the neural networks.
From the experiment, we can observe little improvement in the result of decompilation with the use of attention.
Unfortunately, the results were not applicable for practical use.

@@

In this thesis, we apply neural networks to the problem of decompilation to generate human-friendly pseudocodes. 
This idea is originated from \citet{Motoneta}, in which they used recurrent neural networks, in particular sequence-to-sequence model, for decompilation.
However, their model does not seem to be implemented some features for improving the quality of decompilation.
Therefore, we implemented the attention mechanism, which is introduced in \citet{dot_attention}, for improving the quality.
Additionally, the sequence-to-sequence model, which they used for the modeling, ignores the structure of a C language code.
Therefore, we also implemented the sequence-to-tree model, which is suggested in \citet{Seq2Tree}, for generating structured C language source code.

In the experiment, we tested three models, the basic sequence-to-sequence model, the sequence-to-sequence model with attention and the sequence-to-tree model.
As the result of the experiment, the attention mechanism showed little improvement in the performance, while the sequence-to-tree does not.
Unfortunately, the results of the all models seemed to be not applicable for practical use.
The results of decompilation for short C source code fragment were not so bad. 
Nevertheless, the results for long C codes, which are considered to appear in practical use, were mostly failed.

@@

\chapter{Related Works}
In this chapter, we review related works.

@@

\citet{Motoneta} used the seqence-to-sequence model, which is a popular technique in the machine translation.
The seqence-to-sequence model is also evaluated in this thesis.

\citet{genetic_decompiler} applied genetic programming to decompilation.
They prepared a big corpus of C source codes, then they repeatedly mutated C codes to reduce the difference of the target binary and the compilation result of a C code.

@@

\section{Improvement of decompilation results}
In this thesis, we do not try to recover the function name and variable name. 
However, for hand-decompilation, sometimes the name can be estimated from the functionality of the variable.
For example, \citet{hand_decompilation} used the Boomerang decompiler and repeatedly fixed variable types and names of the decompilation results 
to recover the source code of a software.

@@

\citet{name_recover_from_decompile_result} tackled this work. 
In that paper, they tried to estimate the name of a variable which appears in the decompilation result of a compiled C program.

This kind of name estimation also appears in the context of deobfuscation.
\citep{deobfsucation_matome}
% これ雑にreferenceしたけどあんま意味ないよなぁ...?
Deobfuscation is a reverse process of obfuscation.
Obfuscation is a technique to modify the program source code so that the code becomes hard to understand for humans. 
Obfuscation is used for hiding the behavior of a program whose source code has to be distributed.
For example, in the web application, a JavaScript code of a program should be distributed in order to be executed on client web browsers.
Such JavaScript code is minified in order to protect the program from cracking. 
All comments, newlines, indents are removed and variable names are substituted to automatically generated meaningless names.

\citet{JSNaughty} applied statistical machine translation for this task. 
They calculated the future value of a variable from its functionality, then 
sought the most suitable name from the candidates generated from the corpus.

@@

There are many applications of neural networks for program synthesis. They are summarized by \citet{deep_programming_matome}.

In this thesis, we applied sequence-to-tree model suggested by \citet{Seq2Tree}, 
because the input of the decompiler is a sequence of machine codes and
the output of the decompiler is an AST.

For another example, \citet{coffeescript_to_javascript} tried to mutually translate from JavaScript to CoffeeScript, and its inverse.
In that case, both JavaScript codes and CoffeeScript codes are tree structured, so they suggested tree-to-tree model.

\chapter{Problem Settings and Methods}

@@

A compiler is a program which compiles a source code written in language X to another language Y. 
% 文献 Usually, X is a high-level language and Y is relatively a low-level language. 
For example, a GNU C compiler translates the C language to the x64 assembly language, a javac compiler translates Java to a Java virtual machine code.
% This process is usually 
% 文献
% Especially, 
% This reverse process is usually insufficient or impossible.
A decompiler is a program which aims to reverse the process of a compiler, retrieving a source code of the high-level language X from a source code of a low-level language Y. 
When X is a high-level language and Y is relatively a low-level language, 
this reverse process becomes insufficient or impossible
\citep{hex_rays,decompile_hard_java}.
% for instance X is the C language and Y is the x64 assembly language 
This is because many information, such as variable names or function names, are lost when compiling.
Additionally, the program structure described in the high level-language is lost too. 
For example, a simple \texttt{for} statement can be represented by the combination of a \texttt{while} statement and an \texttt{if} statement, or the combination of \texttt{if} statement and \texttt{goto} statement. 
So a decompiler cannot distinguish the difference in their representations, 
although in the real situation \texttt{for} statement are often used and \texttt{goto} statement are rarely used.

@@

By utilizing a corpus for source codes, a statistical method might have the potential for finding patterns automatically, without enumerating each patterns.
Therefore, we tried to apply some statistical methods, which are described in the next section.

@@

% 189
In this thesis, we try to apply statistical approach, specifically SMT (statistical machine translation) technique, for decompilation.
SMT is the task which tries to translate one sentence $sx$ written in one language into the sentence $sy$ written in another language.

The decompilation problem is formulated as follows. 
Suppose that there is a compiler $ f $ which converts a source code $ sx $ written in a programming language 
to a source code $ sy $  written in another programming language Y with probability $p(sy|sx)$.
In addition, suppose that the human-generated source code $ sx $ follows the probability distribution $ p(sx) $.
In this situation, the decompilation probability distribution $ p(sx|sy) $ in proportion to $ p(sy|sx)p(sx) $ according to the Bayesian low, 
and the $\argmax_{sx} p(sx|sy)$ is considered as the best human-intelligible decompilation result for the code $sy$.

In the experiment, we collected source codes from open source projects as the random sampling from the distribution $ p(sx) $, 
and compiled them to generate the data pairs $ {\{(sx_k,sy_k)\}}_k $.
Then the data pairs are used for training the SMT models to estimate $p(sx|sy)$, and testing the performance of the models.

@@

% (TODO)LSTM,Seq2seqの説明、その後Attentionについて説明、さらにその後seq2treeについて説明する。
\subsection{RNN and LSTM}
The RNN (Recurrent Neural Network) is a neural network for dealing with sequence data.
The structure of the RNN is the sequence of the same node units, as shown in figre~\ref{fig:encoder} and figre~\ref{fig:decoder}. 
The LSTM (Long Short-Term Memory) is a unit of Recurrent Neural Network. 
It was first introduced by \citet{first_LSTM}, for overcoming the vanishing gradient problem and exploding gradient problem.
In short, LSTM unit receives two input vectors, 
the one is the internal state of the LSTM, and the other is the input for the LSTM unit, 
and LSTM emits one output vectors, which represents the next internal state of the LSTM.

@@

The sequence-to-sequence (sequence to sequence) model is the model for SMT task, introduced by \citet{seq2seq}.
% The combination of RNN and LSTM is applied to SMT task in, as the essential unit of the 
Sequence-to-sequence model is composed of two recurrent neural networks, the encoder network and the decoder network.

@@

Figure~\ref{fig:encoder} shows the structure of the encoder network.  The purpose of the encoder network is to summarize the input sequence $sx$, whose length varies depending on the sequence, into one fixed length vector.
This process is carried out as follows. 
First, $sx$ is tokenized to sequence $sx_{1} \dots sx_{n}$, and each token $sx_{k}$ is embedded to its vector representation $vx_{k}$. 
Then, the initial LSTM internal state $enc_0$ is initialized randomly, 
and $ enc_{1} \dots enc_{n} $ are calculated in accordance with the recursive formula $ enc_{k} = lstm_{enc}(enc_{k-1},vx_{k}) $.
Finally, the last internal state $ enc_{n}$ is returned as the reduced representation of the input sequence $sx$.

Figure~\ref{fig:decoder} shows the structure of the decoder network.  The purpose of the decoder network is to estimate the distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $ for each k. 
The training process is carried out as follows. 
First, $sy$ is tokenized, embedded to a vector representation $vy_{1} \dots vy_{m}$ in the same way as embedding $sx$.
% prepended the token $ EOS $ for indicating the end of the sequence, then 
Then, the first internal state $ dec_{1}$ is calculated from the start vector $vy_{begin}$ and encoded input $ enc_{n} = dec_{0} $ in accordance with the formula  $ LSTM_{dec}(enc_{n},y_{begin}) $.
Then, $ dec_{2} \dots dec_{m+1} $ are calculated in accordance with the recursive formula $ dec_{k+1} = lstm_{dec}(dec_{k},vy_{k}) $.
Each decoder internal state $ dec_{k} $ is passed to the linear layer $ W_{dec} $ and then softmaxed to generate a probability vector $ py_{k} $.
The probability vector $ py_{k} $ is intended to represent the probability distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
Thus the network is trained to reduce the difference between predicted distribution $ py_{k} $ and actual distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
In the experiment, we choiced cross-entropy of $ py_{k} $ and $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $ for loss function.

@@

First is the Bidirectional Recurrent Neural Networks (BRNN) is the architecture introduced by \citet{BiRNN}.
In our experiment, this architecture is used for the encoder network as follows.
In addition to the internal state $ enc_{k} $, the reverse internal state $ enc_{k}^{\gets} $ is also calculated 
in accordance with the recursive formula $ enc_{k}^{\gets} = lstm_{enc}^{\gets}(enc_{k+1}^{\gets},vx_{k}) $.
Then both of the internal states $ enc_{n} $ and $ enc_{1}^{\gets} $ are used for decoder network.

Second is the multi-layer recurrent neural networks, introduced by \citet{multi_layer}.
They reported that, feeding the output sequence of the LSTM network to another LSTM network, like stacking the network layers, will improve the accuracy of the prediction.
This architecture is called as stacked recurrent neural networks or multi-layer recurrent neural networks.

@@

The attention module is first developed by \citet{attention_paper}.
We implemented the global general attention, which is introduced by \citet{dot_attention}, for improving the quality of decompilation.
In the following, we will describe about the global general attention.

As above mentioned, the pure sequence-to-sequence model uses the output of decoder network $ dec_{k} $ for estimating 
the distribution $ p(sy_k|sx,[sy_1,\dots,sy_{k-1}]) $.
In the sequence-to-sequence model with attention, 
a context vector $ ctx_{k} $ is calculalted from $ dec_{k} $ and the internal states of encoder network $ enc_{1} \dots enc_{n} $, 
then the distribution is estimated from the output vector $ dec_{k} $ and context vector $ ctx_{k} $.
The context vector $ ctx_{k} $ is calculated as follows.
For each state $ enc_{l} $, the importance of the state $ a_{l} $ is calculated as $ a_{l} = score(dec_{k},enc_{l}) $,
then the $ ctx_{k} $ is calculated as $ ctx_{k} = \sum_{l=1}^{n} \frac{\exp{a_{l}}}{\sum_{l=1}^{n}\exp{a_{l}}} enc{l} $, 
as the weighted average of $ enc_{1} \dots enc_{n} $.
% Multiple score functions are introduced by \cite{dot_attention}, and we choiced 
In the paper by \citet{dot_attention}, various kinds of score function is suggested.
We used general score function, which is calculated as $ score(dec_{k},enc_{l}) = dec_{k}^T W enc_{l} $ ,
where W is the parameter matrix.

@@

\subsection{Sequence-to-Tree}
% 184 words
Compare to natural languages, programming languages obeys rigid grammars. % wich are usually context free grammers.
Therefore, it is reasonable to consider the grammar of the programming language when generating programs.
Sequence-to-Tree is a model for generating tree-structured output from sequence input. 
This model is introduced by \citet{Seq2Tree} for generating a program from a program specification written in the natural language. 
This task is similar to our task, which aims to generate programs from the sequence of the assembly code, so we implemented this model in the experiment.

Now, we will describe about the sequence-2-tree model.
sequence-to-tree model is composed of two modules, the encoder network and the decoder network.
The encoder network of sequence-2-tree model is the same as the encoder network of sequence-to-tree model. It summarize the input sequence into one fixed length vector.

@@

The decoder network is a network for generating tree structured output. 
A tree-structure is representable by the sequence of the actions which expands a nonterminal symbol of the constructing tree.
For example, the tree shown in figure~\ref{fig:ast_zu} is generated from the sequence $ [act_1, \dots act_{12}] $, 
where action $ act_k $ expands the left-most nonterminal symbol in the tree generated from the action sequence $ [act_1, \dots act_{k-1}] $. 
Suppose the tree $ty$ is generated by the action sequence $ [act_1, \dots act_m] $, 
then the tree-generating probability $ p(ty|sx) $ is decomposed to $ \prod_{k=1}^n p(act_k|sx,[act_1, \dots act_{k-1}]) $. 
The sequence-to-sequence model is now applicable for generating tree structure, by embeding each action to representing vector $av_k$ and 
calculating $ lstm_{dec}(dec_{k},av_{k}) $ sequentially for estimating $p(act_{k+1}|sx,[act_1, \dots act_{k}]) $.

@@

In that paper, they additionally tries to use the information which is peculiar to the tree structure. 
For expanding the node $d$, they also use the information of node type $ nf_{d} $ and the parent action $p_{d}$ from which $d$ is generated.
Then they calculated $ lstm_{dec}(dec_{k},[av_{k}; nfv_{d}; pv_{d}]) $ instead of $ lstm_{dec}(dec_{k},av_{k}) $, 
where $nfv_{d},pv_{d}$ are the vector embedding of $nf_{d},p_{d}$.

@@

\chapter{Experiments}
In this chapter, we explain the detail of the data preparation and implementation.
The implementation source codes are available at \url{https://github.com/satos---jp/neural_decompiler}.

\section{Data Collection}
% 219 words
We collected source codes by crawling GitHub \citep{github}, cloning about 900 repositories with C language topic, 
and recursively searched each folder to enumerate C language source code.
From each source code, we generated pairs of C source code fragment and x64 assembly code corresponds to the C fragment.
Some example data pairs are shown in figure~\ref{fig:pairsoffragments}.

@@

The pairs were generated as follows. Let S0 be an original C source code.
\begin{enumerate}
\item Remove all preprocessor, such as \texttt{\#include}, in S0 to generate preprocessed C code S1. 
\item Tokenize S1 and rearrange them in S2 so that there is exactly one token for each line. 
\item 
Compile S2 by the GNU C compiler with no optimization option (-O0) and debug option to get assembly code S3. 
With debug option, S3 contains information about the source code position which each assembly operations are originated from.
\item Assemble S3 to object file S4 and disassemble S4 to get information eliminated assembly code S5.
\item 
Parse S2 with clang to get AST of the source code. 
And for each subtree of AST, we can get C fragment source code which corresponds to the subtree, 
and we can get assembly code corresponds to the C fragment by employing S5 combine with S3.
\end{enumerate}

@@

Second way is the tokenized sequence of assembly codes. 
The disassembled opcodes and operands in S5 are split to symbols, and concatenated these sequences to one sequence, then we modified followings.
\begin{itemize}
\item For each end of the operations, the \texttt{ENDLINE} symbol is added to indicate the end of operation.
\item 
The relative access of code address, such as \texttt{call} or \texttt{jump} operations, 
are relabeled like $\alpha$-conversion for making code position independent and for reducing the vocabulary size.
\item The big integer values are substituted. 
In detail, the integer in the range from -255 to 256 are preserved as it is, and the other integers are substituted with the \texttt{\_\_HEX\_\_} symbol.
This also reduces the vocabulary size.
\end{itemize}

We choosed second way for the experiment.
This is because, in our case, the assembly code is generated from common C compiler, therefore the assembly code can be disassembled uniquely.
For example, \citet{disasm_obfuscate} proposed a obfuscation technique, which deceives the disassembler to generate wrong disassembly result.
In that case, the first way might be better than the second way.

\subsection{C Code Preprocess}
% 232 words
The C source codes were parsed to AST by a Clang compiler \citep{clang} and its Python bindings together with pycparser \citep{pycparser}, 
because all of the tools are individually not sufficient for generating dataset.
Then, the AST is converted to input data by following modifications.

\begin{itemize}
\item The variable-ary nodes and optional nodes are converted. 
For example, the \texttt{CompoundStatement} node has variable number of statements. 
We converted such list of nodes to the chains of concatenation \texttt{CONS} and termination \texttt{NIL}, like a list representation in functional language.
Another example is the \texttt{if} statement, whose \texttt{else} clause are sometimes missing. 
We converted such optional node to the alternative of \texttt{SOME} and \texttt{NONE}, like a optional data structure in functional language.
\item The variable names, function names, label names, or type names were renamed like $\alpha$-conversion.
\item There are four type of literals in C language. 
The integer literal less than 256 is preserved, and the literal greater than 256 is clipped to 256.
The string, float and character literals are converted to \texttt{\_\_LITERAL\_\_} symbol.

Modifying the node to indicate the value in the assembly code corresponding to the literal is also conceivable way,
but we did not implement this because of its technical difficulty, so this is left to future work.
\end{itemize}

@@

\section{Model Configuration}
The training model was implemented with Chainer version 4.5.0 \citep{chainer}.
We implemented three models, Seq2seq, Seq2seqAtt, and Seq2tree.
